---
title: "Project 1"
author: "Michael Costello"
format: html
editor: source
---

In this project, we are creating a series of functions to request data from a Census Bureau API, build robust controls for that function, and then apply the function in several ways, including to append multiple outputs together and also to summarize the data received.

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(httr)
library(jsonlite)
library(janitor)
#install.packages("cli")
library(cli)
```
# Part  1 - Data Processing

I wanted to start off by ensuring I understood the basics of how the API functioned, so I explicitly created a API URL to practice pulling the data down.
```{r}
trial_run <- httr::GET("https://api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION&for=DIVISION:5&key=541a25ef175ec8d1abf1a4b9bb864ac88cf70e54")
  parsed_data <- fromJSON(rawToChar(trial_run$content))
  api_result <- parsed_data |>
      janitor::row_to_names(row_number = 1) |>
      as_tibble() |>
      type_convert()
```
Once I was successful at that task, I turned to finding and downloading the dataset labels. Again, I decided to make sure I could get the data for one year, and then generalized my function.

## Parsing Variable Values (i.e. labels)
Getting one year first.
```{r}
variable_json <- httr::GET("https://api.census.gov/data/2021/acs/acs1/pums/variables.json")

parsed_labels <- fromJSON(rawToChar(variable_json$content))
subset_labels <- parsed_labels$variables[c("AGEP","GASP","GRPIP","JWAP","JWDP", "JWMNP","PWGTP","FER","HHL","SCH","SCHL","SEX","ST","REGION","DIVISION")]
saveRDS(subset_labels, file = "data/variable_descriptions.rds")
```
Here I create a loop, carefully trying to control mis-aligned variables, such as STATE, which was called ST before 2023. I then had the loop save one dataset for year year.

```{r}
years <- 2021:2024
for (yr in years) {
  #Create some strings that I'll need to make the loops run correctly
  api_str <- paste0("https://api.census.gov/data/",yr,"/acs/acs1/pums/variables.json")
  save_str <- paste0("data/variable_descriptions_",yr,".rds")
  
  #Actually collect the data from Census
  variable_json <- httr::GET(api_str)

  #Parse and subset the lists
  parsed_labels <- fromJSON(rawToChar(variable_json$content))
  if(yr <= 2022) {
    subset_labels <- parsed_labels$variables[c("AGEP","GASP","GRPIP","JWAP","JWDP", "JWMNP","PWGTP","FER","HHL","SCH","SCHL","SEX","ST","REGION","DIVISION")]
    #rename ST to STATE to make all more comparable later.
    names(subset_labels)[13] <- "STATE" 
  }
  else subset_labels <- parsed_labels$variables[c("AGEP","GASP","GRPIP","JWAP","JWDP", "JWMNP","PWGTP","FER","HHL","SCH","SCHL","SEX","STATE","REGION","DIVISION")]
  
  #Save the output
  saveRDS(subset_labels, file = save_str)
}
```

At this point I was ready to start writing parts of my actual function. I read through the project parameters, but did not do a sufficient job of planning out how I would break the function into several smaller functions. That would probably have made debugging easier, as well as some aspects of testing and enhancing the code. Instead, I tried to tackle the project in a linear fasion, modeled off the order of the bullet points in the task outline. I built my header and parsed several of the input parameters, including some hard codes (like the beginning of the api's web address). 

```{r}
#| eval: false
helper <- function(year = 2024, vars = c(), restrictions = c()) {

  #Generate all the input parameters:
    address <- "https://api.census.gov/data/"
    vars_allowed <- c("AGEP", "GASP", "GRPIP", "JWAP", "JWDP", "JWMNP", "JWMNP", "PWGTP","FER","HHL","SCH","SCHL","SEX","ST","STATE","REGION","DIVISION")
    
	#Create separate lists of geo, str, and num vars for further checking
      numeric_vars <- c("AGEP","GASP","GRPIP","JWAP","JWDP","JWMNP")
      string_vars  <- c("FER","HHL","SCH","SCHL","SEX")
      geography_vars<- c("ST","STATE","REGION","DIVISION")
```

I then started handling some of the basic errors I know I'd be addressing. First I dealt with the problem of someone trying to reqest multiple years, a year outside of our range (2021-2024), or not supplying variables (ie creating defaults). I decided not to simply apply the defaults in the function header because of several interactions they may have with restriction parameters and other potential problems later on.

```{r}
#| eval: false
  #Check singularity and validity of the year
  if (length(year) != 1) {
    stop("Slow down One Direction, the argument 'year' must be a single value, not a vector or list.")
  }
  if (!year %in% c(2021, 2022, 2023, 2024)) {
    stop("Hey *NSync, the argument 'year' must be ONE of the following: 2021, 2022, 2023, 2024.")
  }

  #Check the requested variables, supply defaults if none requested.
  if (length(vars) == 0) {
    final_request_vars <- "AGEP,PWGTP,SEX"
  } 
```
I then went to work on the myrid permutations and ramification of if the user did request variables, including if those requested variables were not on the list of requestable variables. Included in this step was where I forced the weighting variable `PWGTP` into my string of requesed variables. This chunk begins with an `else` block continued from the `if` block above.

```{r}
#| eval: false

else {
      #Check to make sure all requested vars are allowed.
      if (!all(vars %in% vars_allowed)) {
        stop("Yo, New Kids on the Block, you chose variables that are not in our list. Please restrict to numeric variables: AGEP, GASP, GRPIP, JWAP, JWDP, JWMNP, and to categorical varailes: FER, HHL, SCH, SCHL, SEX.")
      }

      #Create separate lists of the requested variables for each type
       req_num_vars <- paste(vars[vars %in% numeric_vars], collapse = ",")
       req_str_vars <- paste(vars[vars %in% string_vars], collapse = ",")
       req_geo_vars <- paste(vars[vars %in% geography_vars], collapse = ",")
       
       #Push out defaults, in case nothing is listed (in a sub category)
       if (nchar(req_num_vars) == 0) req_num_vars <- "AGEP"
       if (nchar(req_str_vars) == 0) req_str_vars <- "SEX"
       if (nchar(req_geo_vars) == 0) {
         if (year <= 2022) req_geo_vars <- "ST"
         else req_geo_vars <- "STATE"
       } else {
         #########################################
         ###### Strip State out of req_geo_vars  #
         #########################################
       }
       #Glue the requested variable strings all together
       final_request_vars <- paste(req_num_vars,req_str_vars,req_geo_vars,"PWGTP", sep = ",")
  }
```
I wanted to, during this last step, remove STATE from a list of reqested variables because it would be used later, but that code was conflicting with other bits and I just put in a placeholder to return to later.































































